---
title: "Network Kernal Density Estimation (NKDE)"
format: html
date-modified: "last-modified"
execute:  
  warning: false
  freeze: true
  message: false 
  eval: true 
  echo: true 
  fig-retine: 3
date: "`r Sys.Date()`"
---

### 1 Getting Started

```{r}
pacman::p_load(arrow, dplyr, lubridate, ggplot2, tidyverse, sf, sfdep, sp, spNetwork, spdep, raster, spatstat, tmap, readxl, plotly)
```

```{r}
osm_basemap <- tm_basemap(server = "OpenStreetMap.HOT")
imagery_basemap <- tm_basemap(server = "Esri.WorldImagery")
```

### 2. Importing Data

#### 2.1 Asaptial Data

```{r}
hk_census <- read_excel("data/aspatial/hkcensus.xlsx")
```

#### 2.2 Geospatial Data

#### 2.2.1 18 Districts in Hong Kong

```{r}
district_18 <- st_read(dsn = "data/geospatial/hk_18Districts/",
                       layer = "HKDistrict18" )
```

```{r}
sf_district_18 <- district_18 %>% st_transform(crs = 2326)
```

```{r}
st_crs(sf_district_18)
```

```{r}
tmap_mode("plot")
d18_map <- tm_shape(sf_district_18) +
  tm_fill(col = "ENAME", title = "District", legend.show = FALSE) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_text("ID", size = 0.5, col = "black") +
  tm_layout(frame = FALSE)

d18_map
```

#### 2.2.2 Cycling Points in Hong Kong

```{r}
cp <- read_csv("data/aspatial/hkrecyclepoints.csv")
```

Change the geographic location

```{r}
cp_sf <- st_as_sf(cp, 
                  coords = c("lgt","lat"), 
                  crs = 4326) %>% 
  st_transform(crs= 2326)
```

```{r}
summary(cp_sf)
```

```{r}
cp_sf_1 <- cp_sf %>%
  mutate(district_id = toupper(str_replace_all(district_id, "_", " ")))
```

```{r}
recycling_bins <- subset(cp_sf_1, legend == "Recycling Bins at Public Place")
recycling_spots <- subset(cp_sf_1, legend == "Recycling Spots")
private_collection_points <- subset(cp_sf_1, legend == "Private Collection Points (e.g. housing estates, shopping centres)")
ngo_collection_points <- subset(cp_sf_1, legend == "NGO Collection Points")
recycling_stations <- subset(cp_sf_1, legend == "Recycling Stations/Recycling Stores")
street_corner_recycling_shops <- subset(cp_sf_1, legend == "Street Corner Recycling Shops")
smart_bins <- subset(cp_sf_1, legend == "Smart Bin")
```

```{r}
recycling_spots_cp <- st_join(sf_district_18, recycling_spots)
ngo_cp <- st_join(sf_district_18, ngo_collection_points)
pcp_joined_data <- st_join(sf_district_18, private_collection_points)
recycling_bins_cp <- st_join(sf_district_18, recycling_bins)
recycling_stations_cp <- st_join(sf_district_18, recycling_stations)
street_corner_cp <- st_join(sf_district_18, street_corner_recycling_shops)
smart_bins_cp <- st_join(sf_district_18, smart_bins)
```

```{r}
private_collection_points_by_district <- pcp_joined_data %>%
  group_by(ENAME) %>%
  summarize(total_pcp = n())

ngo_cp_by_district <- ngo_cp %>%
  group_by(ENAME) %>%
  summarize(total_ngo_cp = n())

recycling_spots_by_district <- recycling_spots_cp %>%
  group_by(ENAME) %>%
  summarize(total_recycling_spots = n())

recycling_bins_by_district <- recycling_bins_cp %>%
  group_by(ENAME) %>%
  summarize(total_recycling_bins = n())

recycling_stations_by_district <- recycling_stations_cp %>%
  group_by(ENAME) %>%
  summarize(total_recycling_stations = n())

street_corner_shops_by_district <- street_corner_cp %>%
  group_by(ENAME) %>%
  summarize(total_street_corner = n())

smart_bins_by_district <- smart_bins_cp %>%
  group_by(ENAME) %>%
  summarize(total_smart_bins = n())
```

```{r}

```

#### 2.3 Road Data in Hong Kong

```{r}
#| eval: false
road_data <- st_read(dsn = "data/geospatial/china-latest-free.shp", 
                     layer = "gis_osm_roads_free_1")
```

Transform the Data into Hong Kong Projection System

```{r}
#| eval: false
road_data_2326 <- st_transform(road_data, 2326)
```

```{r}
#| eval: false
roads_in_hk <- st_intersection(road_data_2326, sf_district_18)
```

```{r}
#| eval: false
write_rds(roads_in_hk, "data/rds/sf_roads_in_hk.rds")
```

```{r}
roads_in_hk <- read_rds("data/rds/sf_roads_in_hk.rds")
```

### 3. Visualizing Recycling Points in Hong Kong

```{r}
pcp_map <- tm_shape(private_collection_points_by_district) +
  tm_fill(col = "total_pcp") +
  tm_borders() +
  tm_layout(legend.show = TRUE, main.title = "Distribution of Private Collection Points by District",main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()

```

```{r}
ngo_cp_map <- tm_shape(ngo_cp_by_district) +
  tm_fill(col = "total_ngo_cp") +
  tm_borders() +
  tm_layout(legend.show = TRUE, main.title = "Distribution of NGO Points by District", main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
recycling_spots_map <- tm_shape(recycling_spots_by_district) +
  tm_fill(col = "total_recycling_spots") +
  tm_borders() +
  tm_layout(legend.show = TRUE, main.title = "Distribution of Recycling Spots by District",main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
recycling_bins_map <- tm_shape(recycling_bins_by_district) +
  tm_fill(col = "total_recycling_bins") +
  tm_borders() +
  tm_layout(legend.show = TRUE, main.title = "Distribution of Recycling Bins by District", main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
recycling_stations_map <- tm_shape(recycling_stations_by_district) +
  tm_fill(col = "total_recycling_stations") +
  tm_borders() +
  tm_layout(legend.show = TRUE, main.title = "Distribution of Recycling Stations by District", main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
street_corner_shops_map <- tm_shape(street_corner_shops_by_district) +
  tm_fill(col = "total_street_corner") +
  tm_borders() +
  tm_layout(legend.show = TRUE,main.title = "Distribution of Street Corner Shops by District", main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
smart_bin_map <- tm_shape(smart_bins_by_district) +
  tm_fill(col = "total_smart_bins") +
  tm_borders() +
  tm_layout(legend.show = TRUE,main.title = "Distribution of Smart Bins by District",main.title.position = "center", main.title.size = 0.75)+
  tm_scale_bar()
```

```{r}
tmap_arrange(pcp_map,recycling_spots_map, recycling_bins_map,ngo_cp_map, recycling_stations_map, street_corner_shops_map, smart_bin_map, ncol=2, nrow = 4)
```

Since we want to conduct a more accurate analysis on the recycling points, We will focus on the recycling facilities type with a higher number of data - Private Collection Points and Recycling Bins.

#### 3.1 Private Collection Points

```{r}
wm_q_pcp <- private_collection_points_by_district %>%
  mutate(nb = st_contiguity(geometry, queen = TRUE),
         wt = st_weights(nb,
                         style = "W",
                         allow_zero = TRUE),
         .before = 1) 
wm_q_pcp
```

```{r}
set.seed(1234)
global_moran_perm(wm_q_pcp$total_pcp,
                       wm_q_pcp$nb,
                       wm_q_pcp$wt,
                  zero.policy = TRUE,
                  nsim = 999)
```

```{r}
lisa_pcp <- wm_q_pcp %>% 
  mutate(local_moran = local_moran(
    total_pcp, nb, wt, zero.policy = TRUE, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

```{r}
lisa_sig_pcp <- lisa_pcp  %>%
  filter(p_ii < 0.05)

tmap_mode("plot")
tm_shape(lisa_pcp) +
  tm_polygons() +
  tm_borders(alpha = 0.7) +
tm_shape(lisa_pcp) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

#### 3.2 Recycling Bins

```{r}
wm_q_rbins <- recycling_bins_by_district %>%
  mutate(nb = st_contiguity(geometry, queen = TRUE),
         wt = st_weights(nb,
                         style = "W",
                         allow_zero = TRUE),
         .before = 1) 
wm_q_rbins
```

```{r}
set.seed(1234)
global_moran_perm(wm_q_rbins$total_recycling_bins,
                       wm_q_rbins$nb,
                       wm_q_rbins$wt,
                  zero.policy = TRUE,
                  nsim = 999)
```

```{r}
lisa_rbins <- wm_q_rbins %>% 
  mutate(local_moran = local_moran(
    total_recycling_bins, nb, wt, zero.policy = TRUE, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

```{r}
lisa_sig_rbins <- lisa_rbins  %>%
  filter(p_ii < 0.05)

tmap_mode("plot")
tm_shape(lisa_rbins) +
  tm_polygons() +
  tm_borders(alpha = 0.7) +
tm_shape(lisa_rbins) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

Considering the Hot-Spot Analysis on the Private Collection Points and Public Recycling bins, they are maininly locate far from the residential area. Therefore, we decide to refer to the census data in Hong Kong.

```{r}
ggplot(hk_census, aes(x = DNAME, y = POPULATION)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  geom_text(aes(label = POPULATION), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Population by District",
       x = "District Name",
       y = "Population") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Since Sha Tin has the most population, we will focus on Sha Tin district. Moreover, we want to choose another district with both commercial and residential area. Therefore, we will be analyzing on Yau Tsim Mong Area as well.

### 4. Sha Tin District

```{r}
shatin <- sf_district_18 %>%
  filter(ENAME == "SHA TIN")
```

```{r}
plot(st_geometry((shatin)))
```

```{r}
roads_in_shatin <- st_intersection(roads_in_hk, shatin)
```

```{r}
#| eval: false
write_rds(roads_in_shatin, "data/rds/roads_in_shatin.rds")
```

```{r}
roads_in_shatin <- read_rds("data/rds/roads_in_shatin.rds")
```

```{r}
#| eval: false
private_cp_shatin <- st_intersection(private_collection_points, shatin)
```

```{r}
#| eval: false
public_cp_shatin <- st_intersection(recycling_bins, shatin)
```

```{r}
#| eval: false
write_rds(private_cp_shatin, "data/rds/private_cp_shatin.rds")
```

```{r}
#| eval: false
write_rds(public_cp_shatin, "data/rds/public_cp_shatin.rds")
```

```{r}
private_cp_shatin <- read_rds("data/rds/private_cp_shatin.rds")
```

```{r}
public_cp_shatin <- read_rds("data/rds/public_cp_shatin.rds")
```

```{r}
plot(st_geometry(private_cp_shatin))
```

```{r}
plot(st_geometry(public_cp_shatin))
```

```{r}
unique_types <- unique(st_geometry_type(roads_in_shatin))
unique_types
```

```{r}
roads_in_shatin <- roads_in_shatin %>%
  filter(st_geometry_type(geometry) %in% c("LINESTRING", "MULTILINESTRING"))

```

```{r}
if ("LINESTRING" %in% unique_types) {
  roads_in_shatin <- st_cast(roads_in_shatin, "LINESTRING")
} else {
  # handle the case when no linestrings are found
  stop("No linestrings found in roads_in_shatin")
}
```

```{r}
tmap_mode('plot')
tm_shape(roads_in_shatin, geometry_type = "lines") + 
  tm_lines()
```

```{r}
tmap_mode('plot')

# Plot roads
tm_shape(roads_in_shatin, geometry_type = "lines") + 
  tm_lines(lwd = 1, col = "blue") +
  
  # Plot Punggol area boundary
  tm_shape(shatin) +
  tm_borders() +
  
  # Plot Grab origin data
  tm_shape(private_cp_shatin) +
  tm_dots()
```

```{r}
roads_lines_shatin <- roads_in_shatin[st_geometry_type(roads_in_shatin) == "LINESTRING", ]

# Apply lixelize_lines with mindist
lixels_shatin <- lixelize_lines(roads_lines_shatin,5000, mindist = 2500)
```

```{r}
samples_shatin <- lines_center(roads_lines_shatin)
```

#### 4.1 Private Collection Points in Sha Tin

```{r}
densities_shatin_private <- nkde(roads_lines_shatin, 
                  events = private_cp_shatin,
                  w = rep(1,nrow(private_cp_shatin)),
                  samples = samples_shatin,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

```{r}
samples_shatin$density_private <- densities_shatin_private
lixels_shatin$density_private <- densities_shatin_private
```

```{r}
# rescaling to help the mapping
samples_shatin$density_private <- samples_shatin$density_private*1000
lixels_shatin$density_private <- lixels_shatin$density_private*1000
```

```{r}
tmap_mode('view')

shatin_density_private <- osm_basemap+
tm_shape(lixels_shatin)+
  tm_lines(col="density_private")+
tm_shape(private_cp_shatin)+
  tm_dots()

shatin_density_private
 
```

#### 4.2 Public Recycling Bins in Sha Tin

```{r}
densities_shatin_public <- nkde(roads_lines_shatin, 
                  events = public_cp_shatin,
                  w = rep(1,nrow(public_cp_shatin)),
                  samples = samples_shatin,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

```{r}
samples_shatin$density_public <- densities_shatin_public
lixels_shatin$density_public <- densities_shatin_public
```

```{r}
# rescaling to help the mapping
samples_shatin$density_public <- samples_shatin$density_public*1000
lixels_shatin$density_public <- lixels_shatin$density_public*1000
```

```{r}
tmap_mode('view')


shatin_density_public <- osm_basemap+
tm_shape(lixels_shatin)+
  tm_lines(col="density_public")+
tm_shape(public_cp_shatin)+
  tm_dots()

shatin_density_public

```

### 5. Yau Tsim Mong Area

```{r}
ytm <- sf_district_18 %>%
  filter(ENAME == "YAU TSIM MONG")
```

```{r}
plot(st_geometry((ytm)))
```

```{r}
#| eval: false
roads_in_ytm <- st_intersection(roads_in_hk, ytm)
```

```{r}
#| eval: false
write_rds(roads_in_ytm, "data/rds/roads_in_ytm.rds")
```

```{r}
roads_in_ytm <- read_rds("data/rds/roads_in_ytm.rds")
```

```{r}
#| eval: false
private_cp_ytm <- st_intersection(private_collection_points, ytm)
```

```{r}
#| eval: false
public_cp_ytm <- st_intersection(recycling_bins, ytm)
```

```{r}
#| eval: false
write_rds(private_cp_ytm, "data/rds/private_cp_ytm.rds")
```

```{r}
#| eval: false
write_rds(public_cp_ytm, "data/rds/public_cp_ytm.rds")
```

```{r}
private_cp_ytm <- read_rds("data/rds/private_cp_ytm.rds")
```

```{r}
public_cp_ytm <- read_rds("data/rds/public_cp_ytm.rds")
```

```{r}
plot(st_geometry(private_cp_ytm))
```

```{r}
plot(st_geometry(public_cp_ytm))
```

```{r}
unique_types <- unique(st_geometry_type(roads_in_ytm))
unique_types
```

```{r}
roads_in_ytm <- roads_in_ytm %>%
  filter(st_geometry_type(geometry) %in% c("LINESTRING", "MULTILINESTRING"))

```

```{r}
if ("LINESTRING" %in% unique_types) {
  roads_in_ytm <- st_cast(roads_in_ytm, "LINESTRING")
} else {
  # handle the case when no linestrings are found
  stop("No linestrings found in roads_in_ytm")
}
```

```{r}
#| eval: false
tmap_mode('plot')
tm_shape(roads_in_ytm, geometry_type = "lines") + 
  tm_lines()
```

#### 5.1 Private Collection Poitns in YTM

```{r}
#| eval: false
tmap_mode('plot')

# Plot roads
tm_shape(roads_in_ytm, geometry_type = "lines") + 
  tm_lines(lwd = 1, col = "blue") +
  
  tm_shape(ytm) +
  tm_borders() +

  tm_shape(private_cp_ytm) +
  tm_dots()
```

```{r}
roads_lines_ytm <- roads_in_ytm[st_geometry_type(roads_in_ytm) == "LINESTRING", ]

# Apply lixelize_lines with mindist
lixels_ytm <- lixelize_lines(roads_lines_ytm,5000, mindist = 2500)
```

```{r}
samples_ytm <- lines_center(lixels_ytm)
```

```{r}
densities_ytm_private <- nkde(roads_lines_ytm, 
                  events = private_cp_ytm,
                  w = rep(1,nrow(private_cp_ytm)),
                  samples = samples_ytm,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

```{r}
samples_ytm$density_private <- densities_ytm_private
lixels_ytm$density_private <- densities_ytm_private
```

```{r}
# rescaling to help the mapping
samples_ytm$density_private <- samples_ytm$density_private*1000
lixels_ytm$density_private <- lixels_ytm$density_private*1000
```

```{r}
tmap_mode('view')

ytm_density_private <- osm_basemap + 
  tm_shape(lixels_ytm)+
  tm_lines(col="density_private")+
tm_shape(private_cp_ytm)+
  tm_dots()

ytm_density_private
```

#### 5.2 Public Recycling Bins in YTM

```{r}
densities_ytm_public <- nkde(roads_lines_ytm, 
                  events = public_cp_ytm,
                  w = rep(1,nrow(public_cp_ytm)),
                  samples = samples_ytm,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

```{r}
samples_ytm$density_public <- densities_ytm_public
lixels_ytm$density_public <- densities_ytm_public
```

```{r}
# rescaling to help the mapping
samples_ytm$density_public <- samples_ytm$density_public*1000
lixels_ytm$density_public <- lixels_ytm$density_public*1000
```

```{r}
tmap_mode('view')

ytm_density_public <-osm_basemap + 
tm_shape(lixels_ytm)+
  tm_lines(col="density_public")+
tm_shape(private_cp_ytm)+
  tm_dots()

ytm_density_public
```
